---
title: "FINAL PROJECT Q1"
author: "Sixue Cheng"
date: "12/18/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(mice)
library(tidyverse)
library(leaps)
setwd("~/Desktop/math 185/final project")
Dataset <- read.csv("kread.csv")
```

```{r}
sum(is.na(Dataset$R.SCORE))
datasets =na.omit(Dataset)
sum(is.na(Dataset$L.SCORE))
dataset =na.omit(Dataset)
```

(1)
```{r}
inp1 = dataset %>% group_by(INP = 1)
inp2 = dataset %>% group_by(INP = 0)
r_inp1 = inp1$R.SCORE
r_inp2 = inp2$R.SCORE
l_inp1 = inp1$L.SCORE
l_inp2 = inp2$L.SCORE
t.test(r_inp1,r_inp2)
t.test(l_inp1,l_inp2)
```
Since p_value in R_score T-test and L_score T-test of students in these two programs is 1, so there has a significant difference in terms of the reading ability with 2 groups.

(2)
I use best subset selection:
```{r}
bss.full = regsubsets(DIEBLS~ dataset$R.SCORE + dataset$L.SCORE + dataset$FARM, dataset, nvmax = 3)
bss.summary = summary(bss.full)
which.min(bss.summary$bic)
(bss.summary$which)
```
So by best subset selection, 2 variables should be included, and the variables should be R-score and L score.

(3) I separate the dataset into a training set and a testing set, and use test error to determine the adequacy of model:
```{r}
dt = sort(sample(nrow(dataset), nrow(dataset)*.2))
test<-dataset[dt,]
train <- dataset[-dt,]
fit_train = lm(DIEBLS~R.SCORE + L.SCORE, data = train)
summary(fit_train)
D_predict = predict(fit_train, newdata = test)
test_error= mean((D_predict - test$DIEBLS)^2)
test_avg = mean(test$DIEBLS)
tss = sum((test_avg - test$DIEBLS)^2)
rsq = (tss - test_error) / tss
rsq
```
Since R^2 for using L-score and R-score as predictors is 0.9863, so the model is adequate.

(4) we can use similar case to predict the missing value of the data. And in R there is a package called "Mice", we can use functions in this package to predict the missing values.
```{r}
Dataset$R.SCORE=as.factor(Dataset$R.SCORE)
md.pattern(Dataset)
Dataset$L.SCORE=as.factor(Dataset$L.SCORE)
md.pattern(Dataset)
mice_imputes = mice(Dataset)
Imputed_data=complete(mice_imputes)
Imputed_data
dim(Imputed_data)
```

(5) We can use variance test on prediction with FARM and prediction without FARM:
```{r}
DATA <- mutate(dataset, FARM = (FARM == "N")+0)
head(DATA)
fit.w = lm(DIEBLS~ R.SCORE + L.SCORE + FARM, data = DATA)
predict_w = predict(fit.w, newdata = DATA)
fit.wo = lm(DIEBLS~ DATA$R.SCORE + DATA$L.SCORE, data = DATA)
predict_wo = predict(fit.wo, newdata = DATA)
var.test(predict_w, predict_wo, alternative = "two.sided")
```
Since p-value of variance test between prediction with FARM and prediction without FARM is 1, so there is no statistical difference if including FARM as a covariate. So L-score and R-score are already a good model in predicting DIEBLS.
